### 2016-08-12 latex/paper.tex
Change refs to Bibtex

Despite its implications for wheel strength, the buckling problem has never received a rigorous treatment to our knowledge. Jobst Brandt alludes to buckling in his practical manual for wheelbuilders\cite{Brandt}:

@@

He did not discuss the problem further, but suggested that wheel failure commonly occurs due to a loss of lateral stability caused by spoke buckling. Pippard and Francis\cite{Pip2} derived a model for lateral stiffness based on the elastic foundation model, but did not discuss stability and neglected any effects of spoke tension.

Flexural-torsional buckling of rings can be treated as a special case of buckling of arches, where the included angle is allowed to go to $2\pi$. Timoshenko and Gere\cite{Tim} gave a formula for the critical load for a ring with doubly-symmetric cross-section subjected to radial loads. The theory of flexural-torsional buckling of monosymmetric arches (bicycle rims have only one plane of symmetry) was broadly formalized by Trahair and Papangelis\cite{Tra} using the virtual work approach to derive the equilibrium and stability equations. Their theory has been extended to treat arches restrained by continuous\cite{Brad1} or discrete\cite{Brad2} elastic supports and elastic end restraints\cite{Guo}.

The problem of the prestressed bicycle wheel is unique for a number of reasons. First, the buckling load is internal to the structure. Second, the spokes act both as elastic restraints resisting buckling and as prestressing elements causing buckling. Third, the lateral, radial, and torsional restraining actions of the spokes are coupled---i.e. lateral deflection at a spoke may produce lateral, radial, and torsional reactions. These considerations extend to other structural systems. Large observation wheels such as the London Eye\cite{Mann} and the Singapore Flyer\cite{All} resemble large bicycle wheels and achieve lateral stability due to the bracing angle of prestressed cables, and must be designed against flexural-torsional buckling.

@@

% \bibliographystyle{acm}
% \bibliography{biblio}
\begin{thebibliography}{6}
% \begin{thebibliography}{66} % use this for more than 9 references
\bibitem{Brandt} J.~Brandt, 
\textit{The Bicycle Wheel},
Avocet, Palo Alto, (1993).

\bibitem{Pip2} A.J.S.~Pippard, W.E.~Francis, 
``The Stresses in a Wire Wheel under Side Loads on the Rim'',
\textit{Philos. Mag.} \textbf{14} (1932), pp. 436-445.

\bibitem{Tim} S.P.~Timoshenko, J.M.~Gere, 
\textit{Theory of Elastic Stability}, 
McGraw-Hill (1961).

\bibitem{Tra} N.S.~Trahair, J.P.~Papangelis, 
``Flexural-torsional buckling of monosymmetric arches'', 
\textit{J. Struct. Eng.} \textbf{113} (1987), pp. 2271-2288

\bibitem{Brad1} Y.~-L. Pi and M.~A. Bradford, 
``Elastic flexural-torsional buckling of continuously restrained arches'',
\textit{Int. J. Solids Struct.} \textbf{39} (2002), pp. 2299-2322.

\bibitem{Brad2} Y.~-L. Pi and M.~A. Bradford, 
``Elastic flexural-torsional buckling of discretely restrained arches'',
\textit{J. Struct. Eng.} \textbf{39} (2002), pp. 2299-2322.

\bibitem{Guo} Y.L.~Guo, S.Y.~Zhao, C.~You, Y.L.~Pi, 
``Out-of-Plane Elastic Buckling of Circular Arches with Elastic End Restraints'', 
\textit{J. Struct. Eng.} \textbf{140} (9) (2014)

\bibitem{BandD} C.~J. Burgoyne and R. Dilmaghanian,
Bicycle Wheel as Prestressed Structure, 
\textit{J. Eng. Mech.} \textbf{119} (1993), pp. 439-455.

\bibitem{Mann} A.P.~Mann, N.~Thompson, M.~Smits, 
``Building the British Airways London Eye'', 
\textit{Proc. Inst. Civ. Eng. - Civ. Eng.} \textbf{144} (2002), pp. 60-72

\bibitem{All} A.~Allsop, P.~Dallard, B.~McNiven, 
``The Singapore Flyer and Design of Giant Observation Wheels'', 
\textit{Struct. Eng. Int.-IABSE} \textbf{19}(1), (2009), pp. 12-16

\end{thebibliography}

>>>>

Despite its implications for wheel strength, the buckling problem has never received a rigorous treatment to our knowledge. Jobst Brandt alludes to buckling in his practical manual for wheelbuilders\cite{Brandt1993c}:

@@

He did not discuss the problem further, but suggested that wheel failure commonly occurs due to a loss of lateral stability caused by spoke buckling. Pippard and Francis\cite{Pippard1932d} derived a model for lateral stiffness based on the elastic foundation model, but did not discuss stability and neglected any effects of spoke tension.

Flexural-torsional buckling of rings can be treated as a special case of buckling of arches, where the included angle is allowed to go to $2\pi$. Timoshenko and Gere\cite{Timoshenko1961a} gave a formula for the critical load for a ring with doubly-symmetric cross-section subjected to radial loads. The theory of flexural-torsional buckling of monosymmetric arches (bicycle rims have only one plane of symmetry) was broadly formalized by Trahair and Papangelis\cite{Trahair1987b} using the virtual work approach to derive the equilibrium and stability equations. Their theory has been extended to treat arches restrained by continuous\cite{Pi2002b} or discrete\cite{Bradford2002d} elastic supports and elastic end restraints\cite{Guo2014b}.

The problem of the prestressed bicycle wheel is unique for a number of reasons. First, the buckling load is internal to the structure. Second, the spokes act both as elastic restraints resisting buckling and as prestressing elements causing buckling. Third, the lateral, radial, and torsional restraining actions of the spokes are coupled---i.e. lateral deflection at a spoke may produce lateral, radial, and torsional reactions. These considerations extend to other structural systems. Large observation wheels such as the London Eye\cite{Mann2001a} and the Singapore Flyer\cite{Allsop2009a} resemble large bicycle wheels and achieve lateral stability due to the bracing angle of prestressed cables, and must be designed against flexural-torsional buckling.

@@

\bibliographystyle{acm}
\bibliography{C:/Users/matt/OneDrive/Documents/Research/bibtex/Papers-BMD2016.bib}
